---
title: "Background Curve Fitting"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

For each row (individual), this fits a curve of the form:

$$y = 2^{bx} - 1$$

using two points:

- Point 1: (`blank_time`, `blank_rate_abs`)
- Point 2: (`total_time`, `background_rate_abs`)

This curve shape automatically passes through (0, 0) for any value of `b`,
which is why it's used here: many rows have a blank point right at the
origin (rate below detection limit at time 0), and this form handles that
naturally without forcing the whole curve flat.

Since there is only one free parameter (`b`) but two data points, `b` is
estimated by nonlinear least squares (`nls`), minimizing the squared error
across both points rather than solving them exactly.

The fitted curve for each row is then used to calculate the predicted
background rate at `rmr_time` and `mmr_time`.

**Rows with any missing (NA) required value are skipped** and get `NA` for
`b`, `rmr_background_rate_abs`, and `mmr_background_rate_abs`.

## Setup

```{r libraries}
library(dplyr)
library(readr)
```

## Load data

```{r load-data}
infile <- "background_curve_data.csv"
df <- read_csv(infile, show_col_types = FALSE)
```

## Fitting function

```{r fit-function}
# Fits y = 2^(b*x) - 1 to the two points (x1,y1) and (x2,y2) via nls,
# then returns the fitted b along with predictions at rmr_time and mmr_time.
# Returns a named list: b, rmr_pred, mmr_pred (all NA if inputs are missing
# or the fit fails).
fit_background_curve <- function(x1, y1, x2, y2, rmr_x, mmr_x) {

  required <- c(x1, y1, x2, y2, rmr_x, mmr_x)
  if (any(is.na(required))) {
    return(list(b = NA_real_, rmr_pred = NA_real_, mmr_pred = NA_real_))
  }

  # Initial guess for b, based on the background point alone
  ratio <- y2 + 1
  b0 <- if (x2 != 0 && ratio > 0) log2(ratio) / x2 else 0.0001

  fit <- tryCatch({
    nls(
      y ~ 2^(b * x) - 1,
      data  = data.frame(x = c(x1, x2), y = c(y1, y2)),
      start = list(b = b0)
    )
  }, error = function(e) NULL)

  if (is.null(fit)) {
    return(list(b = NA_real_, rmr_pred = NA_real_, mmr_pred = NA_real_))
  }

  b <- coef(fit)[["b"]]
  list(
    b = b,
    rmr_pred = 2^(b * rmr_x) - 1,
    mmr_pred = 2^(b * mmr_x) - 1
  )
}
```

## Apply to every row

```{r apply-fit}
results <- df %>%
  rowwise() %>%
  mutate(
    .fit = list(fit_background_curve(blank_time, blank_rate_abs,
                                      total_time, background_rate_abs,
                                      rmr_time, mmr_time)),
    b = .fit$b,
    rmr_background_rate_abs = .fit$rmr_pred,
    mmr_background_rate_abs = .fit$mmr_pred
  ) %>%
  ungroup() %>%
  select(-.fit)
```

## Summary

```{r summary}
cat("Total rows:              ", nrow(results), "\n")
cat("Rows with a fitted curve:", sum(!is.na(results$b)), "\n")
cat("Rows skipped (NA):       ", sum(is.na(results$b)), "\n")
```

## Sanity-check plot (one example row)

```{r example-plot}
example_row <- results %>% filter(!is.na(b)) %>% slice(1)

x_seq <- seq(0, max(example_row$mmr_time, example_row$rmr_time, example_row$total_time) * 1.05, length.out = 200)
y_seq <- 2^(example_row$b * x_seq) - 1

plot(x_seq, y_seq, type = "l", col = "steelblue", lwd = 2,
     xlab = "Time", ylab = "Rate (abs)",
     main = paste("Fitted curve - fish#", example_row$`fish#`))
points(c(example_row$blank_time, example_row$total_time),
       c(example_row$blank_rate_abs, example_row$background_rate_abs),
       pch = 19, col = "black")
points(c(example_row$rmr_time, example_row$mmr_time),
       c(example_row$rmr_background_rate_abs, example_row$mmr_background_rate_abs),
       pch = 4, col = "red", lwd = 2)
legend("topleft", legend = c("Fitted curve", "Input points (blank/background)", "Predicted (rmr/mmr)"),
       col = c("steelblue", "black", "red"), pch = c(NA, 19, 4), lty = c(1, NA, NA), lwd = c(2, NA, 2))
```

## Write output CSV

```{r write-output}
write_csv(results, "calculated_background_curve.csv")
```
