---
title: "Respirometry Analysis "
author: "Hayden Simm"
date: "2026-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading packages
```{r}
  library(tidyverse)
  library(dplyr)
  library(janitor)
  library(respR)
  library(readr)

```

creating the dataframeS
```{r}
suppressWarnings(
  suppressMessages({
#GROUP1
df_killifish_respirometry_acclimation_group1 <- read_delim("../DATA/KILLIFISH_RESPIROMETRY_ACCLIMATION_GROUP1.txt")
df_background_respirometry_acclimation_group1 <- read_delim("../DATA/BACKGROUND_RESPIROMETRY_ACCLIMATION_GROUP1.txt")
df_killifish_respirometry_opposite_group1 <- read_delim("../DATA/KILLIFISH_RESPIROMETRY_OPPOSITE_GROUP1.txt")
df_background_respirometry_opposite_group1 <- read_delim("../DATA/BACKGROUND_RESPIROMETRY_OPPOSITE_GROUP1.txt")
  })
)
```

Creating a list of all data frames, preserving the original names 
```{r}
dfs_group1 <- mget(c(
  "df_killifish_respirometry_acclimation_group1",
  "df_background_respirometry_acclimation_group1",
  "df_killifish_respirometry_opposite_group1",
  "df_background_respirometry_opposite_group1"
))
```

Function for removing rows that include device information
```{r}
remove_rows <- function(df, marker = "#--- Measurement Data --------------------------------------------------------------------") {
  # Find the row containing the marker
  start_row <- which(apply(df, 1, function(x) any(x == marker)))
  
  # Keep only rows below the marker
  df <- df[(start_row + 1):nrow(df), ]
  
  return(df)
}
```
*note from the author* Alternatively, I attempted to use a function to select columns by the name in the pyroscience software but unfortunately some runs contained NA where the following titles should have been. Following this I have renamed the columns as they should appear below.
select_columns <- function(df) {
  df %>%
    select(
    " dt (s) [A Ch.1 Main]",
    "Oxygen (%air sat.) [A Ch.1 Main]",
    "Oxygen (%air sat.) [A Ch.2 Main]",
    "Oxygen (%air sat.) [A Ch.3 Main]",
    "Oxygen (%air sat.) [A Ch.4 Main]",
    "Oxygen (%air sat.) [B Ch.1 Main]",
    "Oxygen (%air sat.) [B Ch.2 Main]",
    "Oxygen (%air sat.) [B Ch.3 Main]",
    "Oxygen (%air sat.) [B Ch.4 Main]"
    ) %>%
    janitor::clean_names(case = "snake") %>%
    mutate(across(everything(), as.numeric)) %>%
    drop_na()
  }

Renaming columns by what they could/should be in the raw data
```{r}
rename_and_clean <- function(df) {
  
  # Check that there are 9 columns
  if (ncol(df) != 9) {
    stop("Dataframe does not have 9 columns.")
  }
  
  colnames(df) <- c(
    "dt (s) [A Ch.1 Main]",
    "Oxygen (%air sat.) [A Ch.1 Main]",
    "Oxygen (%air sat.) [A Ch.2 Main]",
    "Oxygen (%air sat.) [A Ch.3 Main]",
    "Oxygen (%air sat.) [A Ch.4 Main]",
    "Oxygen (%air sat.) [B Ch.1 Main]",
    "Oxygen (%air sat.) [B Ch.2 Main]",
    "Oxygen (%air sat.) [B Ch.3 Main]",
    "Oxygen (%air sat.) [B Ch.4 Main]"
  )
  
  df %>%
    janitor::clean_names(case = "snake") %>%
    mutate(across(everything(), as.numeric)) %>%
    drop_na()
}
```


```{r}
process_functions <- function(df) {
  df %>%
    remove_rows() %>%
    select(3,4,22,40,58,81,99,117,135) %>%
    rename_and_clean()
}

dfs_processed_group1 <- lapply(dfs_group1, process_functions)
```

This is an optional step I recommend removing the original list because of its large size
```{r}
rm(dfs_group1,
   df_killifish_respirometry_acclimation_group1,
   df_killifish_respirometry_opposite_group1,
   df_background_respirometry_acclimation_group1,
   df_background_respirometry_opposite_group1
)

```

#RMR Acclimation
```{r}
fish7_acclimation <- dfs_processed_group1[["df_killifish_respirometry_acclimation_group1"]] %>%
  select("dt_s_a_ch_1_main", "oxygen_percent_air_sat_a_ch_4_main") %>%
  inspect(time = 1, oxygen = 2, plot = FALSE) %>%
  subset_data(from = 20000, to = 30000, by = "time") %>% #subset to after MMR and middle of night 
  auto_rate.int(start = 600, #time that flush pump is on, in seconds
                wait = 200, 
                measure = 280, #wait + measure times should add to time that flush pump is off
                method = "linear", width = 200, by = "time") %>%
  convert_rate(oxy.unit = '%Air', time.unit = 's', output.unit = 'umol/min/g', 
               volume = 0.175, S = 11, t = 10, mass = 0.008597) %>%
  #plot(pos = ) %>%
  #summary(pos = ) #%>% this is to specifically look at specific replicates
  select_rate(method = "rsq", n = c(0.9,1)) %>%
  select_rate(method = "lowest_percentile", n = 0.1) %>%
  mean() %>% #this is your final value
  summary(export = TRUE)
```

#RMR Opposite
```{r}
fish7_opposite <- dfs_processed_group1[["df_killifish_respirometry_opposite_group1"]] %>%
  select("dt_s_a_ch_1_main", "oxygen_percent_air_sat_b_ch_4_main") %>%
  inspect(time = 1, oxygen = 2, plot = FALSE) %>%
  subset_data(from = 20000, to = 30000, by = "time") %>% #subset to after MMR and middle of night 
  auto_rate.int(start = 240, #time that flush pump is on, in seconds
                wait = 80, 
                measure = 100, #wait + measure times should add to time that flush pump is off
                method = "linear", width = 80, by = "time") %>%
  convert_rate(oxy.unit = '%Air', time.unit = 's', output.unit = 'umol/min/g', 
               volume = 0.175, S = 11, t = 24, mass = 0.008597) %>%
  # plot(pos = ) %>%
  # summary(pos = ) %>% #this is to look at specific replicates
  select_rate(method = "rsq", n = c(0.9,1)) %>%
  select_rate(method = "lowest_percentile", n = 0.1) %>%
  mean() %>% #this is your final value
  summary(export = TRUE)
```

if you need a specific position
```{r}
fish6_opposite <- dfs_processed_group1[["df_killifish_respirometry_opposite_group1"]] %>%
  select("dt_s_a_ch_1_main", "oxygen_percent_air_sat_b_ch_3_main") %>%
  inspect(time = 1, oxygen = 2, plot = FALSE) %>%
  subset_data(from = 20000, to = 30000, by = "time") %>%
  auto_rate.int(start = 240, wait = 80, measure = 100,
                method = "linear", width = 80, by = "time") %>%
  convert_rate(oxy.unit = '%Air', time.unit = 's', output.unit = 'umol/min/g',
               volume = 0.175, S = 11, t = 24, mass = 0.005839)

# Step 2: save the pos =  summary as its own object
fish6_opposite <- summary(fish6_opposite, pos = 9, export = TRUE)
```

#MMR AcCLIMATION
```{r}
fish7_acclimation_mmr <- dfs_processed_group1[["df_killifish_respirometry_acclimation_group1"]] %>%
  select("dt_s_a_ch_1_main", "oxygen_percent_air_sat_a_ch_4_main") %>%
  subset_data(from = 350, to = 450, by = "row") %>% #subset to time when MMR was preformed
  inspect(time = 1, oxygen = 2, plot = TRUE) %>%
  auto_rate(method = "linear", width = 0.2, by = "row") %>%
  convert_rate(oxy.unit = '%Air', time.unit = 's', output.unit = 'umol/min/g', 
               volume = 0.175, S = 11, t = 10, mass = 0.008597) %>%
  ## ^adjust volume (in liters), salinity (ppt), temperature (C), mass (in kg)
  select_rate(method = "rsq", n = c(0.9,1)) %>%
  select_rate(method = "rank") %>%
  select_rate(method = "highest", n = 1) %>%
  summary(export = TRUE) 
```

#MMR OPPOSITE
```{r}
fish7_opposite_mmr <- dfs_processed_group1[["df_killifish_respirometry_opposite_group1"]] %>%
  select("dt_s_a_ch_1_main", "oxygen_percent_air_sat_b_ch_4_main") %>%
  subset_data(from = 1050, to = 1200, by = "row") %>% #subset to time when MMR was preformed
  inspect(time = 1, oxygen = 2, plot = TRUE) %>%
  auto_rate(method = "linear", width = 0.2, by = "row") %>%
  #plot(pos = 6) %>% ##to look at different ranks, pick the rank that covers the obvious region
  convert_rate(oxy.unit = '%Air', time.unit = 's', output.unit = 'umol/min/g', 
               volume = 0.175, S = 11, t = 24, mass = 0.008597) %>%
  ## ^adjust volume (in litres), salinity (ppt), temperature (C), mass (in kg)
  select_rate(method = "rsq", n = c(0.9,1)) %>%
  select_rate(method = "rank") %>%
  select_rate(method = "highest", n = 1) %>%
  summary(export = TRUE) 
```

```{r}
group1_data <- bind_rows(
  mget(c("fish4_acclimation_mmr", 
         "fish4_opposite_mmr",
         "fish5_acclimation_mmr", 
         "fish5_opposite_mmr",
         "fish6_acclimation_mmr", 
         "fish6_opposite_mmr",
         "fish7_acclimation_mmr", 
         "fish7_opposite_mmr")),
  .id = "fish_ID"
)
```

```{r}
group1_mmr_data <- bind_rows(
  mget(c("fish4_acclimation_mmr", 
         "fish4_opposite_mmr",
         "fish5_acclimation_mmr", 
         "fish5_opposite_mmr",
         "fish6_acclimation_mmr", 
         "fish6_opposite_mmr",
         "fish7_acclimation_mmr", 
         "fish7_opposite_mmr")),
  .id = "fish_ID"
)
```

```{r}
write.csv(group1_mmr_data, "../ANALYSIS/group1_mmr_data.csv", row.names = FALSE)
```

#Background ACCLIMATION
```{r}
fish4_acclimation_background <- dfs_processed_group1[["df_background_respirometry_acclimation_group1"]] %>%
  select("dt_s_a_ch_1_main", "oxygen_percent_air_sat_a_ch_4_main") %>%
  subset_data(from = 350, to = 450, by = "row") %>% #subset to time when MMR was preformed
  inspect(time = 1, oxygen = 2, plot = TRUE) %>%
  auto_rate(method = "linear", width = 0.2, by = "row") %>%
  convert_rate(oxy.unit = '%Air', time.unit = 's', output.unit = 'umol/min/g', 
               volume = 0.175, S = 11, t = 10, mass = 0.001) %>%
  ## ^adjust volume (in liters), salinity (ppt), temperature (C), mass (in kg)
  select_rate(method = "rsq", n = c(0.9,1)) %>%
  select_rate(method = "rank") %>%
  select_rate(method = "highest", n = 1) %>%
  summary(export = TRUE) 
```

#Background OPPOSITE
```{r}
fish7_opposite_background <- dfs_processed_group1[["df_background_respirometry_opposite_group1"]] %>%
  select("dt_s_a_ch_1_main", "oxygen_percent_air_sat_b_ch_4_main") %>%
  subset_data(from = 1050, to = 1200, by = "row") %>% #subset to time when MMR was preformed
  inspect(time = 1, oxygen = 2, plot = TRUE) %>%
  auto_rate(method = "linear", width = 0.2, by = "row") %>%
  #plot(pos = 6) %>% ##to look at different ranks, pick the rank that covers the obvious region
  convert_rate(oxy.unit = '%Air', time.unit = 's', output.unit = 'umol/min/g', 
               volume = 0.175, S = 11, t = 24, mass = 0.001) %>%
  ## ^adjust volume (in litres), salinity (ppt), temperature (C), mass (in kg)
  select_rate(method = "rsq", n = c(0.9,1)) %>%
  select_rate(method = "rank") %>%
  select_rate(method = "highest", n = 1) %>%
  summary(export = TRUE) 
```

```{r}
group1_background_data <- bind_rows(
  mget(c("fish4_acclimation_background", 
         "fish4_opposite_background",
         "fish5_acclimation_background", 
         "fish5_opposite_background",
         "fish6_acclimation_background", 
         "fish6_opposite_background",
         "fish7_acclimation_background", 
         "fish7_opposite_background")),
  .id = "fish_ID"
)
```

```{r}
group1_background_data %>%
  select(-mass, -rate.m.spec) %>%
  write.csv("../ANALYSIS/group1_background_data.csv", row.names = FALSE)
```